---
title: "Individual 2"
author: "Arlena Wang"
date: "12/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


```{r}
library(tidyverse)
library(readxl)
library(ggplot2)
library(dplyr)
library(ggrepel)
library(lubridate)
library(reshape2)
```

#import the data
```{r}
data <- readxl::read_xlsx('ClimateChange.xlsx', sheet='Data')
country <- readxl::read_xlsx('ClimateChange.xlsx', sheet='Country')
series <- readxl::read_xlsx('ClimateChange.xlsx', sheet='Series')
```

#change the ".." cells in the years to n/a or 0
```{r}
data <- na_if(data, '..')
```

#Questions Investigated
1.) Which country had the most cause/impact towards climate change from 1990-2011?
2.) How does the US gas emissions compare to other countries?
3.) What was the population growth/change over the time period 1990-2011?
4.) Does the amount of gas emissions affect the total population?


**1.) Which country had the most cause/impact towards climate change from 1990-2011?**

**Variable utilized:**
EN.ATM.CO2E.KT = CO2 emissions, total (KtCO2)

```{r}
gas <- data %>% filter(`Series code` == 'EN.ATM.CO2E.KT') %>% 
  summarize(Country_Name =`Country name`, Series_Code =`Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)

gas<-melt(gas, id.vars=c("Country_Name", "Series_Code"))

gas <- gas %>%
  group_by(Country_Name) %>%
  summarize(Country_Name,Series_Code, year=`variable`,gas_emission=`value`)

gas$Series_Code <- NULL

gassignificant <- filter(gas, Country_Name %in% c("China", "Germany", "India", "Iran, Islamic Rep.", "Japan", "Korea, Dem. Rep.", "Russian Federation", "United States"))

ggplot(gassignificant, aes(x=year, y=gas_emission)) + geom_bar(stat="identity") + facet_wrap(~Country_Name, scales="free") + labs(x="Year", y="Gas Emissions", title="Gas Emissions from 1990-2011") + theme(axis.text.x = element_text(angle = 90))
```


**2.) How does the US gas emissions compare to other countries?**



**3.) What was the population growth/change over the time period 1990-2011?**

**Variable utilized:**
SP.POP.GROW=population growth (annual %)

```{r}
pop <- data %>% filter(`Series code` == 'SP.POP.GROW') %>% 
  summarize(`Country name`, `Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)

pop<-melt(pop, id.vars=c("Country name", "Series code"))

pop <- pop %>%
  group_by(`Country name`) %>%
  summarize(Country_Name=`Country name`,`Series code`, year=`variable`,population=`value`)

pop$`Series code` <- NULL
pop$`Country name` <- NULL

popsignificant <- filter(pop, Country_Name %in% c("China", "Germany", "India", "Iran, Islamic Rep.", "Japan", "Korea, Dem. Rep.", "Russian Federation", "United States"))

ggplot(popsignificant, aes(x=year, y=population)) + geom_point() + facet_wrap(~Country_Name) + labs(x="Year", y="Population Growth (annual %)", title="Population Growth from 1990-2011") + theme(axis.text.x = element_text(angle = 90)) 
```

**3.) 4.) Does the amount of gas emissions affect the total population?**

**Variable utilized:**
EN.ATM.CO2E.KT = CO2 emissions, total (KtCO2)
SP.POP.TOTL= total population


```{r}

gas <- data %>% filter(`Series code` == 'EN.ATM.CO2E.KT') %>% 
  summarize(Country_Name =`Country name`, Series_Code =`Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)

gas<-melt(gas, id.vars=c("Country_Name", "Series_Code"))

gas <- gas %>%
  group_by(Country_Name) %>%
  summarize(Country_Name,Series_Code, year=`variable`,gas_emission=`value`)

gas$Series_Code <- NULL

gassignificant <- filter(gas, Country_Name %in% c("China", "Germany", "India", "Iran, Islamic Rep.", "Japan", "Korea, Dem. Rep.", "Russian Federation", "United States"))

totalpop <- data %>% filter(`Series code` == 'SP.POP.TOTL') %>% 
  summarize(`Country name`, `Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)

totalpop<-melt(totalpop, id.vars=c("Country name", "Series code"))

totalpop <- totalpop %>%
  group_by(`Country name`) %>%
  summarize(Country_Name=`Country name`,`Series code`, year=`variable`,Total_population=`value`)

totalpop$`Series code` <- NULL
totalpop$`Country name` <- NULL

totalpopsignificant <- filter(totalpop, Country_Name %in% c("China", "Germany", "India", "Iran, Islamic Rep.", "Japan", "Korea, Dem. Rep.", "Russian Federation", "United States"))

effect <- totalpopsignificant %>%
  inner_join(gassignificant %>%
          select(Country_Name, year, gas_emission),
          by=c("Country_Name", "year"))

ggplot(effect, aes(x=gas_emission, y=Total_population)) + geom_point() + facet_wrap(~Country_Name, scales="free") + labs(x="Gas Emissions (KtCO2)", y="Total Population", title="Total Population Vs. Gas Emissions from 1990-2011") + theme(axis.text.x = element_text(angle = 90))
```


