---
title: "Project Report"
author: "Arlena Wang and Neha Maddali"
date: "12/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Climate Change Dataset Analysis Project Report

**Description of Background**
Climate change refers to long-term shifts in temperatures and weather patterns.  As this is a natural occurring problem, technologies from man-made inventions has driven this phenomenon to become a bigger problem more quickly.  The burning of fossil fuels, such as coal, oil, and gas, for our technologies, produces heat-trapping gases.  The dataset, "Climate Change," investigates many aspects and statistics of countries around the world.  In order to find the countries that have the most impact on climate change, gas emissions, population, environmental factors, and GDP will be explored as these are the main factors that play a role in society which lead to the causes of global warming.

**Questions Raised**
*What was the population growth/change over the time period 1990-2011 in the US? 

*Does the amount of gas emissions affect the total population? 

*How does the US gas emissions compare to gas emissions in other countries? 

*How does GDP affect amount of gas emissions? 

*How does population growth affect GDP? 

*Which country has the most environmental impact from climate change in the time period 1990-2011? 

*Which country had the most cause/impact towards climate change from 1990-2011? 

**Dataset**
The dataset was obtained from the website, [climate change](https://datacatalog.worldbank.org/search/dataset/0040205). This dataset includes information about climate change and was collected from the Climate Change Knowledge Portal and World Development indicators.  The data includes greenhouse gas emissions, energy consumption, and more from the time period 1990-2011. The dataset has three different sheets: Data, Country, Series. Data all the relevant data for climate change collected for the years 1990-2011. Country has a list of all the countries data has been collected on. Series has a summary of the different abbreviations for the types of data (essentially a data code book). It provides explanation and sources to each of the categories of data collected for further research to benefit analysis.

## Data Analysis Exploration

**1.) What was the population growth/change over the time period 1990-2011?**

**Variable utilized:**
SP.POP.GROW=population growth (annual %)

```{r}
pop <- data %>% filter(`Series code` == 'SP.POP.GROW') %>% 
  summarize(`Country name`, `Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)

pop<-melt(pop, id.vars=c("Country name", "Series code"))

pop <- pop %>%
  group_by(`Country name`) %>%
  summarize(Country_Name=`Country name`,`Series code`, year=`variable`,population=`value`)

pop$`Series code` <- NULL
pop$`Country name` <- NULL

popsignificant <- filter(pop, Country_Name %in% c("China", "Germany", "India", "Iran, Islamic Rep.", "Japan", "Korea, Dem. Rep.", "Russian Federation", "United States"))

ggplot(popsignificant, aes(x=year, y=population)) + geom_point() + facet_wrap(~Country_Name) + labs(x="Year", y="Population Growth (annual %)", title="Population Growth from 1990-2011") + theme(axis.text.x = element_text(angle = 90)) 
```

The figure shows a general decrease in population growth for annual percentage as years increase from 1990 to 2011.  While this may be coincidental, it may be due to problems that are caused by climate change.

**2.) Does the amount of gas emissions affect the total population?**

**Variable utilized:**
EN.ATM.CO2E.KT = CO2 emissions, total (KtCO2)
SP.POP.TOTL= total population

```{r}
gas <- data %>% filter(`Series code` == 'EN.ATM.CO2E.KT') %>% 
  summarize(Country_Name =`Country name`, Series_Code =`Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)

gas<-melt(gas, id.vars=c("Country_Name", "Series_Code"))

gas <- gas %>%
  group_by(Country_Name) %>%
  summarize(Country_Name,Series_Code, year=`variable`,gas_emission=`value`)

gas$Series_Code <- NULL

gassignificant <- filter(gas, Country_Name %in% c("China", "Germany", "India", "Iran, Islamic Rep.", "Japan", "Korea, Dem. Rep.", "Russian Federation", "United States"))

totalpop <- data %>% filter(`Series code` == 'SP.POP.TOTL') %>% 
  summarize(`Country name`, `Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)

totalpop<-melt(totalpop, id.vars=c("Country name", "Series code"))

totalpop <- totalpop %>%
  group_by(`Country name`) %>%
  summarize(Country_Name=`Country name`,`Series code`, year=`variable`,Total_population=`value`)

totalpop$`Series code` <- NULL
totalpop$`Country name` <- NULL

totalpopsignificant <- filter(totalpop, Country_Name %in% c("China", "Germany", "India", "Iran, Islamic Rep.", "Japan", "Korea, Dem. Rep.", "Russian Federation", "United States"))

effect <- totalpopsignificant %>%
  inner_join(gassignificant %>%
          select(Country_Name, year, gas_emission),
          by=c("Country_Name", "year"))

ggplot(effect, aes(x=Total_population, y=gas_emission)) + geom_point() + facet_wrap(~Country_Name, scales="free") + labs(x="Gas Emissions (KtCO2)", y="Total Population", title="Gas Emissions Vs. Total Population from 1990-2011") + theme(axis.text.x = element_text(angle = 90))
```

The figure shows a general positive relationship between total population and amount of gas emissions.  The Russian Federation shows a decrease then increase in amount of gas emissions.  Germany had a general decrease between total population and gas emissions.  Germany and the Russian Federation may be trying to significantly reduce their fuel emissions.  The general pattern of an increase in population correlates with an increase in gas emissions for most countries.  This is due to more people causing more activity and fuel consumption which leads to more gas emissions.

**3.) How does GDP affect amount of gas emissions?**

**Variable utilized:**
EN.ATM.CO2E.KT = CO2 emissions, total (KtCO2)
NY.GDP.MKTP.CD = GDP($)
```{r warning=FALSE}
gas <- data %>% filter(`Series code` == 'EN.ATM.CO2E.KT') %>% 
  summarize(Country_Name =`Country name`, Series_Code =`Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)
gas<-melt(gas, id.vars=c("Country_Name", "Series_Code"))
gas <- gas %>%
  group_by(Country_Name) %>%
  summarize(Country_Name,Series_Code, year=`variable`,gas_emission=`value`)
gas$Series_Code <- NULL

gassignificant <- filter(gas, Country_Name %in% c("China", "Germany", "India", 
                                                  "Iran, Islamic Rep.", "Japan", 
                                                  "Korea, Dem. Rep.", "Russian Federation", 
                                                  "United States"))

gdp <- data %>% filter(`Series code` == 'NY.GDP.MKTP.CD') %>% 
  summarize(`Country name`, `Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)
gdp<-melt(gdp, id.vars=c("Country name", "Series code"))
gdp <- gdp %>%
  group_by(`Country name`) %>%
  summarize(Country_Name=`Country name`,`Series code`, year=`variable`,GDP=`value`)
gdp$`Series code` <- NULL
gdp$`Country name` <- NULL

gdpSig <- filter(gdp, Country_Name %in% c("China", "Germany", "India", "Iran, Islamic Rep.", "Japan", "Korea, Dem. Rep.", "Russian Federation", "United States"))

effect <- gdpSig %>%
  inner_join(gassignificant %>%
               select(Country_Name, year, gas_emission),
             by=c("Country_Name", "year"))
effect <- na.omit(effect)                           # Apply na.omit function

#multiple graphs in one figure
ggplot(effect, aes(x=gas_emission, y=GDP)) + geom_point() + facet_wrap(~Country_Name, scales="free") + 
  labs(x="Gas Emissions (KtCO2)", y="GDP", title="GDP Vs. Gas Emissions from 1990-2011") + 
  theme(axis.text.x = element_text(angle = 90)) + geom_smooth(method="lm") 
```
The figure shows a general positive relationship between GDP in dollars and the amount of gas emissions, although there are some outliers. Essentially suggesting that a growing per capita GDP leads to increasing carbon dioxide emissions. Germany is the only country that shows a decreasing trend of these two variables. India, China, and the United States show very interesting trends. China has a large jump in the time frame of 1995-2000. India as well has the large jump between 1999-2001. This could be due to a number of factors such as change in income based on population. And although the United States doesn't show a significantly visible positive correlation, the general trend line would be positive. Overall, economic growth exerts a significant positive effect on greenhouse gases.

**4.) How does population growth affect GDP?**

**Variable utilized:**
<br />
SP.POP.GROW = Population growth (annual %)
<br />
NY.GDP.MKTP.CD = GDP($)
```{r warning=FALSE}
pop <- data %>% filter(`Series code` == 'SP.POP.TOTL') %>% 
  summarize(`Country name`, `Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)

pop<-melt(pop, id.vars=c("Country name", "Series code"))

gdp <- data %>% filter(`Series code` == 'NY.GDP.MKTP.CD') %>% 
  summarize(`Country name`, `Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)

gdp<-melt(gdp, id.vars=c("Country name", "Series code"))

test <- pop %>% inner_join(gdp %>% select("Country name",`Series code`,`variable`,`value`), 
                           by=c("Country name","variable"))
test <- na.omit(test)                           # Apply na.omit function


data1 <- test %>% group_by(`Country name`) %>%
  summarize(`Country name`, "year" = `variable`, "population" = `value.x`, "GDP"=`value.y`)

datSig <- filter(data1, `Country name` %in% c("China", "Germany", "India", "Iran, Islamic Rep.", 
                                            "Japan", "Korea, Dem. Rep.", "Russian Federation", 
                                            "United States"))
ggplot(datSig, aes(x=population, y=GDP, colour = `Country name`, group = `Country name`)) + geom_point()+ geom_line() +labs(x="Population Growth (annual %)", y="GDP ($)", title="GDP Vs. Population from 1990-2011")
```
From the figure, an increase in population has a positive correlation with GDP. There were a few significant jumps in data points for some countries like India, China, Iran/Islamic Rep., and the United States. This could've been caused by some N/A or 0 values where no significant data was recorded for that year. In general, labour is a factor of production and with an increase in the labour force, due to population growth, the total output may increase causing the GDP to increase.

**5.) Which country has the most environmental impact from climate change in the time period 1990-2011?**

**Variable utilized:**
<br />
EN.CLC.MDAT.ZS = Droughts, floods, extreme temps (% pop. avg. 1990-2009)
<br />
ER.H2O.FWTL.ZS = Annual freshwater withdrawals (% of internal resources)
<br />
AG.LND.EL5M.ZS = Land area below 5m (% of land area)

```{r warning=FALSE}
disaster <- data %>% filter(`Series code` == 'EN.CLC.MDAT.ZS') %>% 
  summarize(`Country name`, `Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)
disaster<-melt(disaster, id.vars=c("Country name", "Series code"))

disaster <- filter(disaster, `Country name` %in% c("China", "Germany", "India", 
                                                  "Iran, Islamic Rep.", "Japan", 
                                                  "Korea, Dem. Rep.", "Russian Federation", 
                                                  "United States"))
h2oWithdraw <- data %>% filter(`Series code` == 'ER.H2O.FWTL.ZS') %>% 
  summarize(`Country name`, `Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)
h2oWithdraw<-melt(h2oWithdraw, id.vars=c("Country name", "Series code"))

h2oWithdraw <- filter(h2oWithdraw, `Country name` %in% c("China", "Germany", "India", 
                                                   "Iran, Islamic Rep.", "Japan", 
                                                   "Korea, Dem. Rep.", "Russian Federation", 
                                                   "United States"))
landBelow <- data %>% filter(`Series code` == 'AG.LND.EL5M.ZS') %>% 
  summarize(`Country name`, `Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)
landBelow<-melt(landBelow, id.vars=c("Country name", "Series code"))

landBelow <- filter(landBelow, `Country name` %in% c("China", "Germany", "India", 
                                                         "Iran, Islamic Rep.", "Japan", 
                                                         "Korea, Dem. Rep.", "Russian Federation", "United States"))

impact <- disaster %>% inner_join(h2oWithdraw %>% select("Country name",`Series code`,`variable`,`value`), 
                           by=c("Country name","variable"))
impact <- impact %>% inner_join(landBelow %>% select("Country name",`Series code`,`variable`,`value`), 
                                  by=c("Country name","variable"))
impact <- impact %>% group_by(`Country name`) %>%
  summarize(`Country name`, "year" = `variable`, "Natural Disasters" = `value.x`, 
            "H20 Withdrawal"=`value.y`, "Land Below"=`value`) %>% 
  mutate(`Natural Disasters` = as.numeric(`Natural Disasters`)) %>% 
  mutate(`H20 Withdrawal` = as.numeric(`H20 Withdrawal`)) %>% 
  mutate(`Land Below` = as.numeric(`Land Below`))

impactChina <- filter(impact, `Country name` %in% c("China"))
a <- mean(impactChina$`Natural Disasters`,na.rm=TRUE)
a1 <- mean(impactChina$`H20 Withdrawal`,na.rm=TRUE)
a2 <- mean(impactChina$`Land Below`,na.rm=TRUE)

impactGermany <- filter(impact, `Country name` %in% c("Germany"))
b <- mean(impactGermany$`Natural Disasters`,na.rm=TRUE)
b1 <- mean(impactGermany$`H20 Withdrawal`,na.rm=TRUE)
b2 <- mean(impactGermany$`Land Below`,na.rm=TRUE)

impactIndia <- filter(impact, `Country name` %in% c("India"))
c <- mean(impactIndia$`Natural Disasters`,na.rm=TRUE)
c1 <- mean(impactIndia$`H20 Withdrawal`,na.rm=TRUE)
c2 <- mean(impactIndia$`Land Below`,na.rm=TRUE)

impactUS <- filter(impact, `Country name` %in% c("United States"))
d <- mean(impactUS$`Natural Disasters`,na.rm=TRUE)
d1 <- mean(impactUS$`H20 Withdrawal`,na.rm=TRUE)
d2 <- mean(impactUS$`Land Below`,na.rm=TRUE)

impactJapan <- filter(impact, `Country name` %in% c("Japan"))
e <- mean(impactJapan$`Natural Disasters`,na.rm=TRUE)
e1 <- mean(impactJapan$`H20 Withdrawal`,na.rm=TRUE)
e2 <- mean(impactJapan$`Land Below`,na.rm=TRUE)

impactRussia <- filter(impact, `Country name` %in% c("Russian Federation"))
f <- mean(impactRussia$`Natural Disasters`,na.rm=TRUE)
f1 <- mean(impactRussia$`H20 Withdrawal`,na.rm=TRUE)
f2 <- mean(impactRussia$`Land Below`,na.rm=TRUE)

impactKorea <- filter(impact, `Country name` %in% c("Korea, Dem. Rep."))
g <- mean(impactKorea$`Natural Disasters`,na.rm=TRUE)
g1 <- mean(impactKorea$`H20 Withdrawal`,na.rm=TRUE)
g2 <- mean(impactKorea$`Land Below`,na.rm=TRUE)

impactIran <- filter(impact, `Country name` %in% c("Iran, Islamic Rep."))
h <- mean(impactIran$`Natural Disasters`,na.rm=TRUE)
h1 <- mean(impactIran$`H20 Withdrawal`,na.rm=TRUE)
h2 <- mean(impactIran$`Land Below`,na.rm=TRUE)

country <- c('China','Germany','India','United States','Japan','Russian Federation','Korea, Dem. Rep.','Iran, Islamic Rep.')
avgNaturalDisasters <- c(a,b,c,d,e,f,g,h)
avgH20Withdrawal <- c(a1,b1,c1,d1,e1,f1,g1,h1)
avgLandBelow <- c(a2,b2,c2,d2,e2,f2,g2,h2)

envirImpact <- data.frame(country, avgNaturalDisasters, avgH20Withdrawal, avgLandBelow)
envirImpact <- envirImpact %>% rowwise() %>% 
  mutate(total = sum(c(avgNaturalDisasters, avgH20Withdrawal, avgLandBelow))) %>% 
  summarize(country, avgNaturalDisasters, avgH20Withdrawal, avgLandBelow, total)

lbls <- c('China','Germany','India','United States','Japan','Russian Federation','Korea, Dem. Rep.','Iran, Islamic Rep.')
pct <- round(envirImpact$total/sum(envirImpact$total)*100)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels
pie(envirImpact$total,labels = lbls, col=rainbow(length(lbls)),
    main="Pie Chart of Countries")
```
For this question, the following variables were used: EN.CLC.MDAT.ZS which is droughts, floods, extreme temps (% pop. avg), ER.H2O.FWTL.ZS which is Annual freshwater withdrawals (% of internal resources), and AG.LND.EL5M.ZS which is Land area below 5m (% of land area). After transposing the data for these specific cateogies and calculating a total value of these impacts, the figure shows that Iran/Islamic Rep was the most impacted from climate change. India is the next with highest impact. This data and finding correlates with that of what is currently being researched and reported. Iran has been heavily impacted by climate change. It has reduced their rainfall per annum and the inflow of water into rivers has reduced by 40% which in turn affects agriculture, industrial and drinking water. Similarly, India is experiencing increasing heat waves, droughts, and erratic rainfall events.

**6.) How does the US gas emissions compare to other countries?**

**Variable utilized:**
EN.ATM.CO2E.KT = CO2 emissions, total (KtCO2)

```{r}
gas <- data %>% filter(`Series code` == 'EN.ATM.CO2E.KT') %>% 
  summarize(Country_Name =`Country name`, Series_Code =`Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)

gas<-melt(gas, id.vars=c("Country_Name", "Series_Code"))

gas <- gas %>%
  group_by(Country_Name) %>%
  summarize(Country_Name,Series_Code, year=`variable`,gas_emission=`value`)

gas$Series_Code <- NULL

gassignificant <- filter(gas, Country_Name %in% c("China", "Germany", "India", "Iran, Islamic Rep.", "Japan", "Korea, Dem. Rep.", "Russian Federation", "United States"))

ggplot(gassignificant, aes(x=year, y=gas_emission)) + geom_bar(stat="identity") + facet_wrap(~Country_Name, scales="free") + labs(x="Year", y="Gas Emissions (KtCO2)", title="Gas Emissions from 1990-2011") + theme(axis.text.x = element_text(angle = 90))
```
This figure shows only the most relevant countries with larger quantities of gas emission.  This includes China, Germany, India, Iran, Islamic Republic, Japan, Korea, Democratic Republic, Russian Federation, and the United States.

It seems that US has the second most gas emissions as it also has an increasing value of gas emissions over the years with its highest value at 5461013.74 million metric tons of CO2.  This may be due the amount of transportation, electricity, industry, and agriculture that is present in the US.  There are some fluctuations that can be due to changes in economy, price of fuel, and other factors.  Overall, it follows a similar trend to other countries as there is a positive relationship between year and amount of gas emissions.  

**7.) Which country had the most cause/impact towards climate change from 1990-2011?**

**Variable utilized:**
EN.ATM.CO2E.KT = CO2 emissions, total (KtCO2)

Looking at the same figure above, as climate change is mainly caused by gas emissions, from years 1990 to 2011, China has contributed the most towards climate change based on the provided figure.  When looking at the figure, China has an increasing gas emission value as years pass with the highest gas emission in 2008, at 7031916.21 million metric tons of CO2.  This presents relevant information as China maintains a very large infrastructure development which emits a large amount of CO2 during the refining process.  Many construction-related activities from this cause the CO2 emissions.

**Conclusion**
As previously mentioned, there were quite a few empty values for some of the variables throughout the years. We didn’t want to consider these values as 0 because they could’ve simply not have been any data to collect for that year. More specifically, the major questions we were looking into were which countries caused the most impact towards climate change and which countries were most affected by climate change. From looking at specific variables like population growth, total population, GDP and total CO2 emissions, we found that significant contributors to climate change factors were China and the United States. For looking at variables like droughts, floods, extreme temps, annual freshwater withdrawals, land area below 5m we found that the countries that were most impacted from climate change factors were Iran/Islamic Rep. and India. Other conclusions made were that population growth leads to an increase in GDP which in turn leads to an increase in gas emissions and gas emissions are one of the main causes of climate change. 


**Contribution Statements**
  For this project, the work was split up evenly between Neha Maddali (Individual 1) and Arlena Wang (Individual 2) and an equal amount of contribution was made to the project.  The data was cleaned by both indivual 1 and 2.  Questions 3, 4, and 5 were analyzed by individual 1 while questions 1, 2, 6, and 7 were analyzed by individual 2.  Other work including the presentation was equally completed by both individuals.