---
title: "Project Report"
author: "Arlena Wang"
date: "12/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Climate Change Dataset Analysis Project Report

**Description of Background**
Climate change refers to long-term shifts in temperatures and weather patterns.  As this is a natural occurring problem, technologies from man-made inventions has driven this phenomenon to become a bigger problem more quickly.  The burning of fossil fuels, such as coal, oil, and gas, for our technologies, produces heat-trapping gases.  The dataset, "Climate Change," investigates many aspects and statistics of countries around the world.  In order to find the countries that have the most impact on climate change, gas emissions, population, environmental factors, and GDP will be explored as these are the main factors that play a role in society which lead to the causes of global warming.

**Questions Raised**
*What was the population growth/change over the time period 1990-2011 in the US? 

*Does the amount of gas emissions affect the total population? 

*How does the US gas emissions compare to gas emissions in other countries? 

*How does GDP affect amount of gas emissions? 

*How does population growth affect GDP? 

*Which country has the most environmental impact from climate change in the time period 1990-2011? 

*Which country had the most cause/impact towards climate change from 1990-2011? 

**Dataset**
The dataset was obtained from the website, [climate change](https://datacatalog.worldbank.org/search/dataset/0040205). This dataset includes information about climate change and was collected from the Climate Change Knowledge Portal and World Development indicators.  The data includes greenhouse gas emissions, energy consumption, and more from the time period 1990-2011. The dataset has three different sheets: Data, Country, Series. Data all the relevant data for climate change collected for the years 1990-2011. Country has a list of all the countries data has been collected on. Series has a summary of the different abbreviations for the types of data (essentially a data code book). It provides explanation and sources to each of the categories of data collected for further research to benefit analysis.

## Data Analysis Exploration

**1.) What was the population growth/change over the time period 1990-2011?**

**Variable utilized:**
SP.POP.GROW=population growth (annual %)

```{r}
pop <- data %>% filter(`Series code` == 'SP.POP.GROW') %>% 
  summarize(`Country name`, `Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)

pop<-melt(pop, id.vars=c("Country name", "Series code"))

pop <- pop %>%
  group_by(`Country name`) %>%
  summarize(Country_Name=`Country name`,`Series code`, year=`variable`,population=`value`)

pop$`Series code` <- NULL
pop$`Country name` <- NULL

popsignificant <- filter(pop, Country_Name %in% c("China", "Germany", "India", "Iran, Islamic Rep.", "Japan", "Korea, Dem. Rep.", "Russian Federation", "United States"))

ggplot(popsignificant, aes(x=year, y=population)) + geom_point() + facet_wrap(~Country_Name) + labs(x="Year", y="Population Growth (annual %)", title="Population Growth from 1990-2011") + theme(axis.text.x = element_text(angle = 90)) 
```

The figure shows a general decrease in population growth for annual percentage as years increase from 1990 to 2011.  While this may be coincidental, it may be due to problems that are caused by climate change.

**2.) Does the amount of gas emissions affect the total population?**

**Variable utilized:**
EN.ATM.CO2E.KT = CO2 emissions, total (KtCO2)
SP.POP.TOTL= total population

```{r}
gas <- data %>% filter(`Series code` == 'EN.ATM.CO2E.KT') %>% 
  summarize(Country_Name =`Country name`, Series_Code =`Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)

gas<-melt(gas, id.vars=c("Country_Name", "Series_Code"))

gas <- gas %>%
  group_by(Country_Name) %>%
  summarize(Country_Name,Series_Code, year=`variable`,gas_emission=`value`)

gas$Series_Code <- NULL

gassignificant <- filter(gas, Country_Name %in% c("China", "Germany", "India", "Iran, Islamic Rep.", "Japan", "Korea, Dem. Rep.", "Russian Federation", "United States"))

totalpop <- data %>% filter(`Series code` == 'SP.POP.TOTL') %>% 
  summarize(`Country name`, `Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)

totalpop<-melt(totalpop, id.vars=c("Country name", "Series code"))

totalpop <- totalpop %>%
  group_by(`Country name`) %>%
  summarize(Country_Name=`Country name`,`Series code`, year=`variable`,Total_population=`value`)

totalpop$`Series code` <- NULL
totalpop$`Country name` <- NULL

totalpopsignificant <- filter(totalpop, Country_Name %in% c("China", "Germany", "India", "Iran, Islamic Rep.", "Japan", "Korea, Dem. Rep.", "Russian Federation", "United States"))

effect <- totalpopsignificant %>%
  inner_join(gassignificant %>%
          select(Country_Name, year, gas_emission),
          by=c("Country_Name", "year"))

ggplot(effect, aes(x=Total_population, y=gas_emission)) + geom_point() + facet_wrap(~Country_Name, scales="free") + labs(x="Gas Emissions (KtCO2)", y="Total Population", title="Gas Emissions Vs. Total Population from 1990-2011") + theme(axis.text.x = element_text(angle = 90))
```

The figure shows a general positive relationship between total population and amount of gas emissions.  The Russian Federation shows a decrease then increase in amount of gas emissions.  Germany had a general decrease between total population and gas emissions.  Germany and the Russian Federation may be trying to significantly reduce their fuel emissions.  The general pattern of an increase in population correlates with an increase in gas emissions for most countries.  This is due to more people causing more activity and fuel consumption which leads to more gas emissions.

**3.) How does GDP affect amount of gas emissions?**

**4.) How does population growth affect GDP?**

**5.) Which country has the most environmental impact from climate change in the time period 1990-2011?**

**6.) How does the US gas emissions compare to other countries?**

**Variable utilized:**
EN.ATM.CO2E.KT = CO2 emissions, total (KtCO2)

```{r}
gas <- data %>% filter(`Series code` == 'EN.ATM.CO2E.KT') %>% 
  summarize(Country_Name =`Country name`, Series_Code =`Series code`,`1990`,`1991`,`1992`,`1993`,`1994`,`1995`,`1996`,
            `1997`,`1998`,`1999`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,
            `2008`,`2009`,`2010`,`2011`)

gas<-melt(gas, id.vars=c("Country_Name", "Series_Code"))

gas <- gas %>%
  group_by(Country_Name) %>%
  summarize(Country_Name,Series_Code, year=`variable`,gas_emission=`value`)

gas$Series_Code <- NULL

gassignificant <- filter(gas, Country_Name %in% c("China", "Germany", "India", "Iran, Islamic Rep.", "Japan", "Korea, Dem. Rep.", "Russian Federation", "United States"))

ggplot(gassignificant, aes(x=year, y=gas_emission)) + geom_bar(stat="identity") + facet_wrap(~Country_Name, scales="free") + labs(x="Year", y="Gas Emissions (KtCO2)", title="Gas Emissions from 1990-2011") + theme(axis.text.x = element_text(angle = 90))
```
This figure shows only the most relevant countries with larger quantities of gas emission.  This includes China, Germany, India, Iran, Islamic Republic, Japan, Korea, Democratic Republic, Russian Federation, and the United States.

It seems that US has the second most gas emissions as it also has an increasing value of gas emissions over the years with its highest value at 5461013.74 million metric tons of CO2.  This may be due the amount of transportation, electricity, industry, and agriculture that is present in the US.  There are some fluctuations that can be due to changes in economy, price of fuel, and other factors.  Overall, it follows a similar trend to other countries as there is a positive relationship between year and amount of gas emissions.  

**7.) Which country had the most cause/impact towards climate change from 1990-2011?**

**Variable utilized:**
EN.ATM.CO2E.KT = CO2 emissions, total (KtCO2)

Looking at the same figure above, as climate change is mainly caused by gas emissions, from years 1990 to 2011, China has contributed the most towards climate change based on the provided figure.  When looking at the figure, China has an increasing gas emission value as years pass with the highest gas emission in 2008, at 7031916.21 million metric tons of CO2.  This presents relevant information as China maintains a very large infrastructure development which emits a large amount of CO2 during the refining process.  Many construction-related activities from this cause the CO2 emissions.

**Conclusion**




**Contribution Statements**
  For this project, the work was split up evenly between Neha Maddali (Individual 1) and Arlena Wang (Individual 2) and an equal amount of contribution was made to the project.  The data was cleaned by both indivual 1 and 2.  Questions 3, 4, and 5 were analyzed by individual 1 while questions 1, 2, 6, and 7 were analyzed by individual 2.  Other work including the presentation was equally completed by both individuals.